\documentclass[journal,10pt,twoside]{IEEEtran}

\usepackage{fontspec}
\defaultfontfeatures{Ligatures=TeX}

\usepackage{amsmath,amssymb,mathtools}
\usepackage{booktabs,multirow,array,tabularx}
\usepackage{placeins}
\usepackage{graphicx}
\usepackage{listings,xcolor}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage{cite}

\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  showstringspaces=false,
  frame=single,
  framesep=4pt,
  xleftmargin=4pt,
  columns=fullflexible,
  keepspaces=true,
}

\begin{document}

\title{Reproducing DeSFAM and Deploying It on a Live Kubernetes Cluster with Cilium Tetragon}

\author{Minh \\ KMA --- Specialised Coursework Report (2025--2026)
\thanks{Reproduction of: S.~Zehra et al.,
``DeSFAM: An Adaptive eBPF and AI-Driven Framework for Securing Cloud
Containers in Real Time,'' \textit{IEEE Access}, vol.~13, 2025,
DOI:~10.1109/ACCESS.2025.3592192.}}

\maketitle

\begin{abstract}
We present an independent reproduction and deployment study of
SyscallAD, the syscall anomaly detector at the core of the DeSFAM
framework. On the DongTing dataset, the reported $F_{1}=0.92$ is
reproduced to within sampling noise ($F_{1}=0.932$ for the
VAE+iForest ensemble, $F_{1}=0.963$ for the VAE alone) after
substituting an undocumented normalisation step in the reference
pipeline with a principled quantile-band scaler. Beyond reproduction,
we collect a labelled syscall dataset of $\approx\!16$k events
directly from Cilium Tetragon's gRPC stream on three production-style
container workloads, deploy the trained detector as an in-cluster
gRPC consumer, and evaluate six model variants on three test splits.
Two structural limitations of the published pipeline are surfaced.
First, the DongTing-trained VAE generalises poorly to real container
syscall traffic ($\text{AUC}=0.135$, statistically below random), and
retraining the same architecture on Container data restores
$\text{AUC}=1.0$. Second, the iForest component of the ensemble
contributes negligible discriminative signal under either training
distribution, while supervised baselines (LSTM, 1-D~CNN) on the same
feature space are competitive. Per-window scoring latency is
$9$--$18$~ms on a single CPU core, dominated by feature construction.
We give concrete recommendations for in-environment threshold
calibration and ensemble weighting at deployment time.
\end{abstract}

\input{experience_en}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
